---
title: "Mapping SF Medicines"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    social : menu
    source_code: embed
    vertical_layout: fill
    runtime: shiny
    theme: simplex
 #source:https://testing-apps.shinyapps.io/flexdashboard-highcharter/   
---

```{r}
library(shiny)
library(flexdashboard)
library(tidyverse)
library(tidyr) 
library(maps)
library(wbstats)
library(leaflet)
library(DT)
library(ggplot2)
library(dplyr)
library(VIM)
library(readxl)
library(flextable)
library(plotly)
library(janitor)
library(scales)
library(highcharter)

knitr::opts_chunk$set(message = FALSE)

#Import and organize data

country <- read_excel("C:/Gram project/Datasets/Samples.xlsx")

```
# About

Column 1 {data-width=350}
-----------------------------------------------------------------------

This application is in support with the IDDO's Medicine Quality Research Group 
objective, "Increasing understanding of the prevalence and distribution of 
SF (Substandard and falsified) medical products around the world".
It's aim is to show the distribution of SF products across the world. 

The data in this application are from surveyor, medicine quality tool developed
by the IDDO's IT team.The application is built in R language and application's
layout is produced with the flexdashboard package,charts are produced by
ggplot2, plotly and highchart packages.

```{r}
#Organizing data

country<-country %>% filter(studytype =="Prevalence survey",
            drugtype %in% c("Antibiotic","Antimalarial")) %>% 
  mutate(country= recode(country,"Unknown"="NULL")) %>% 
  mutate(drugstatus= recode(drugstatus,"Unknown"="NULL")) %>% 
  mutate(country = ifelse(country == "NULL", NA, country))

country[country == "NULL"] <- NA

#impute for the missing
newdata <- kNN(country, k=5)
country<-newdata %>% dplyr::select(pubid:studytype)

country<-country %>% filter(tsamples!=-1)
country$qsamples<-as.numeric(country$qsamples)

```

```{r}
Clean_data<-country %>% mutate(country= recode(country,"Côte d'Ivoire"="Ivory Coast")) %>% 
  mutate(status= case_when(drugstatus %in% c("Falsified","Substandard","Poor quality")~"SF",
  TRUE ~ "Good"))

```

# Sidebar {.sidebar}
Select a medicine type to display:
```{r}
medicine_choices <- rev(unique(Clean_data$drugtype))
selectInput("medicine_choice", label = "Display medicine type:",
            choices =medicine_choices, selected = medicine_choices[[1]])
```

Click the __Overall__ tab to view the distribution of substandard and falsified
medicine across the world from all the sampling methods. The __Sampling__ tab 
shows data collected from convenience or random surveys respectively.The __About__
tab defines the objective of the application, the source of data and the tools 
used to develop the application.

Application author: [Lawrence Babu](https://www.tropicalmedicine.ox.ac.uk/team/lawrence-adipo)

Data source : [MQ Surveyor](https://mqdm.iddo.org), University of Oxford, UK


# Overall

Column 2 {data-width=350}
-----------------------------------------------------------------------
### Summary and action items
This report provides summary of data from antibiotics and antimalarial drugs. 

* Number of publications, N=149.
* Survey data points, N=1726, 931 (unique).
* Samples collected 25579.

### Review data

#### Sample status by count

```{r}
tsapls<-country %>% group_by(drugstatus) %>% 
  summarise(Count=sum(qsamples))

#dataset for number of samples
df <- data.frame(drugstatus=c("Falsified","Good quality","Poor quality",
            "Substandard"), ntimes=c(600,20574,3920,485))
df <- as.data.frame(lapply(df, rep, df$ntimes))


df %>% count(drugstatus) %>% 
  adorn_totals() %>% 
  rename("Drugstatus"=drugstatus, "Count"= n) %>% 
  knitr::kable()

```

Column 3 {.tabset data-width=650}
-----------------------------------------------------------------------
### Samples by country
```{r}
#summarise samples per country
all_samples<-Clean_data %>% 
 group_by(country) %>% 
  summarise(Count = n(),
            all_samples=sum(qsamples))

#world data
wdata<-map_data("world") %>%
  filter(region != "Antarctica")

#merge with sample data
merged<-merge(x=wdata,y=all_samples, by.x = "region", 
              by.y = "country", all.x = TRUE) %>% arrange(order)
```

```{r}
renderPlot({
#merged <-merged %>%  filter(samplingtype==input$sampling_choice)

#Plot the map
 ggplot(data = merged, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = all_samples)) + theme_bw() + xlab(NULL) + ylab(NULL) +
  labs(fill='Total samples', caption = "Source:https://mqdm.iddo.org/"
  #title = str_glue("Samples collected from {input$sampling_choice}")
  ) + theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  ) 

})

```
  
### Good samples

```{r}
# Prepare data for quality of samples
Gsamples<-Clean_data %>% filter(status=="Good") %>% 

   group_by(country) %>% 
   summarise(Count = n(),
             gsamples=sum(qsamples),
             tsamples=sum(tsamples),
             sfsamples=(tsamples-gsamples)) %>%
   mutate(prpgs=(gsamples/tsamples)*100,
          prpfs=(sfsamples/tsamples)*100)

```
```{r}

m1<-merge(x=wdata,y=Gsamples, by.x = "region", 
                by.y = "country", all.x = TRUE) %>% arrange(order)

#Map for good samples
renderPlot({
  
#merged <- merged %>%  filter(samplingtype==input$sampling_choice)
ggplot(data = m1, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prpgs)) +
  labs(fill='% Good',caption = "Source:https://mqdm.iddo.org/") + 
    theme_bw() + xlab(NULL) + ylab(NULL) +  theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )

})

```

### SF

```{r}
renderPlot({
  
#merged<- merged %>%  filter(samplingtype==input$sampling_choice)
ggplot(data = m1, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prpfs)) +
  labs(fill='% SF',caption = "Source:https://mqdm.iddo.org/") + 
    theme_bw() + xlab(NULL) + ylab(NULL) +
    theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )
    
 })

```
### SF by year of sample collection

```{r}
renderPlot({
sf <- read_excel("C:/Gram project/Datasets/sf.xlsx")

sf<-sf %>% filter(styear>=2000) %>% mutate(pct=prp/100,
                                           lbl = scales::percent(pct))

p <- ggplot(data=sf, aes(x=styear, y=pct, fill= status))
p+  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(breaks = seq(0, 1, .2), label = percent) +
  geom_text(aes(label =lbl ), 
            size = 3, 
            position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Set2") +
  labs(y = "Percent", 
       fill="Sample status",
       x = "Year of sample collection",
       title = "",caption = "Source:https://mqdm.iddo.org/") +
  theme_classic() + coord_flip() + theme(legend.background = element_rect( 
                                  fill="white", colour="white"))
})

```
# Convenience survey {data-navmenu="Sampling type"}

Column 4 {.tabset data-width=650}
-----------------------------------------------------------------------
```{r}
csamples<-Clean_data %>%filter(samplingtype %in% c("Convenience survey")) %>% 
  
  group_by(country) %>% 
  summarise(count=n(),
            samples=sum(qsamples)) 

```
### Samples by country

```{r}
#world data
wdata<-map_data("world") %>%
  filter(region != "Antarctica")

#merge with sample data
m2<-merge(x=wdata,y=csamples, by.x = "region", 
              by.y = "country", all.x = TRUE) %>% arrange(order)
```

```{r}
renderPlot({
  
#merged <-merged %>%  filter(samplingtype==input$sampling_choice)

#Plot the map
ggplot(data = m2, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = samples)) + theme_bw() + xlab(NULL) + ylab(NULL) +
  labs(fill='Total samples',caption = "Source:https://mqdm.iddo.org/" 
  #title = str_glue("Samples collected from {input$sampling_choice}")
  ) + theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )
  
})
```

### Good samples

```{r}
#  Prepare data for quality of samples
Gsamples<-Clean_data %>% filter(status=="Good",samplingtype=="Convenience survey") %>% 

   group_by(country) %>% 
   summarise(Count = n(),
             gsamples=sum(qsamples),
             tsamples=sum(tsamples),
             sfsamples=(tsamples-gsamples)) %>%
   mutate(prpgs=(gsamples/tsamples)*100,
          prpfs=(sfsamples/tsamples)*100)

  
 m3<-merge(x=wdata,y=Gsamples, by.x = "region", 
                by.y = "country", all.x = TRUE) %>% arrange(order)
```

```{r}
#Map for good samples
renderPlot({
  
#merged <- merged %>%  filter(samplingtype==input$sampling_choice)
ggplot(data = m3, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prpgs)) + theme_bw() + xlab(NULL) + ylab(NULL) +
  labs(fill='% Good',caption = "Source:https://mqdm.iddo.org/") + theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )
})

```

### SF

```{r}
renderPlot({
  
#merged<- merged %>%  filter(samplingtype==input$sampling_choice)
ggplot(data = m3, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prpfs)) + theme_bw() + xlab(NULL) + ylab(NULL) +
  labs(fill='% SF',caption = "Source:https://mqdm.iddo.org/")+ theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )
})

```
### SF by year of sample collection

```{r}

sfyear<-Clean_data %>% filter(samplingtype=="Convenience survey", status=="Good") %>% 
  
  group_by(styear) %>% 
  summarise(count=n (),
            Good=sum(qsamples),
            ysamples=sum(tsamples),
            SF=ysamples-Good)

h <- highchart() %>% 
  hc_xAxis(categories = sfyear$styear) %>% 
  hc_add_series(name = "Good", 
                data = sfyear$Good) %>% 
  hc_add_series(name = "SF", 
                data = sfyear$SF) 

h
```

```{r}

# customize interactive line chart

h <- h %>%
  hc_title(text = "SF by year",
           margin = 20, 
           align = "left",
           style = list(color = "steelblue")) %>% 
  hc_subtitle(text = "1990 to 2020",
              align = "left",
              style = list(color = "#2b908f", 
                           fontWeight = "bold")) %>% 
  hc_credits(enabled = TRUE, # add credits
             text = "Surveyor Data",
             href = "https://mqdm.iddo.org/") %>% 
  hc_legend(align = "left", 
            verticalAlign = "top",
            layout = "vertical", 
            x = 0, 
            y = 100) %>%
  hc_tooltip(crosshairs = TRUE, 
             backgroundColor = "#FCFFC5",
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)

h

```

# Random survey {data-navmenu="Sampling type"}

Column 5 {.tabset data-width=650}
-----------------------------------------------------------------------
```{r}

Rsamples<-Clean_data %>%filter(samplingtype %in% c("Random survey")) %>% 
  
  group_by(country) %>% 
  summarise(count=n(),
            samples=sum(qsamples)) 

```
### Samples by country

```{r}
#world data
wdata<-map_data("world") %>%
  filter(region != "Antarctica")

#merge with sample data
m4<-merge(x=wdata,y=Rsamples, by.x = "region", 
              by.y = "country", all.x = TRUE) %>% arrange(order)
```

```{r}
renderPlot({
  
#merged <-merged %>%  filter(samplingtype==input$sampling_choice)

#Plot the map
ggplot(data = m4, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = samples))+ theme_bw()  + xlab(NULL) + ylab(NULL)+ 
  labs(fill='Total samples',caption = "Source:https://mqdm.iddo.org/" 
  #title = str_glue("Samples collected from {input$sampling_choice}")
  ) + theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )
  
})

```

### Good samples

```{r}
#  Prepare data for quality of samples
Gsamples<-Clean_data %>% filter(status=="Good",samplingtype=="Random survey") %>% 

   group_by(country) %>% 
   summarise(Count = n(),
             gsamples=sum(qsamples),
             tsamples=sum(tsamples),
             sfsamples=(tsamples-gsamples)) %>%
   mutate(prpgs=(gsamples/tsamples)*100,
          prpfs=(sfsamples/tsamples)*100)

 m5<-merge(x=wdata,y=Gsamples, by.x = "region", 
                by.y = "country", all.x = TRUE) %>% arrange(order)
```

```{r}

#Map for good samples
renderPlot({
  
#merged <- merged %>%  filter(samplingtype==input$sampling_choice)
ggplot(data = m5, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prpgs)) + theme_bw() + xlab(NULL) + ylab(NULL)+
  labs(fill='% Good',caption = "Source:https://mqdm.iddo.org/") + theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )

})

```

### SF

```{r}
renderPlot({
  
#merged<- merged %>%  filter(samplingtype==input$sampling_choice)
ggplot(data = m5, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = prpfs)) +  theme_bw() + xlab(NULL) + ylab(NULL) +
  labs(fill='% SF',caption = "Source:https://mqdm.iddo.org/") + theme(
    plot.background = element_blank()
   ,panel.grid.major = element_blank()
   ,panel.grid.minor = element_blank()
   ,panel.border = element_blank()
   ,axis.ticks = element_blank()
   ,axis.text.x = element_blank()
   ,axis.text.y = element_blank()
   ,legend.position = "bottom"
   ,legend.direction = "horizontal"
  )
})

```
### SF by year of sample collection

```{r}

ryear<-Clean_data %>% filter(samplingtype=="Random survey", status=="Good") %>% 
  
  group_by(styear) %>% 
  summarise(count=n (),
            Good=sum(qsamples),
            ysamples=sum(tsamples),
            SF=ysamples-Good)


i <- highchart() %>% 
  hc_xAxis(categories = ryear$styear) %>% 
  hc_add_series(name = "Good", 
                data = ryear$Good) %>% 
  hc_add_series(name = "SF", 
                data = ryear$SF) 

i
```


```{r}
# customize interactive line chart

i <- i %>%
  hc_title(text = "SF by year",
           margin = 20, 
           align = "left",
           style = list(color = "steelblue")) %>% 
  hc_subtitle(text = "1990 to 2020",
              align = "left",
              style = list(color = "#2b908f", 
                           fontWeight = "bold")) %>% 
  hc_credits(enabled = TRUE, # add credits
             text = "Surveyor Data",
             href = "https://mqdm.iddo.org/") %>% 
  hc_legend(align = "left", 
            verticalAlign = "top",
            layout = "vertical", 
            x = 0, 
            y = 100) %>%
  hc_tooltip(crosshairs = TRUE, 
             backgroundColor = "#FCFFC5",
             shared = TRUE, 
             borderWidth = 4) %>% 
  hc_exporting(enabled = TRUE)

i

```


